---
title: "edX - Statistics and R"
author: "Maurizio La Rosa"
date: "13/01/2021"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
# First assessment: Exercises
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Exercise #1

<font size="0.7">0 points possible (ungraded)</font>

What version of R are you using (hint: make sure you download the latest version and then type `version`)? **Please note that this question does not count toward your grade**, but it is important to make sure that you are using the latest version of R. If the answer is not the MOST updated, please just let us know.

```{r 1, echo=FALSE}
print(version)
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Exercise #2

<font size="0.7">1 point possible (graded)</font>

Create a numeric vector containing the numbers 2.23, 3.45, 1.87, 2.11, 7.33, 18.34, 19.23. What is the average of these numbers?

```{r 2, echo=FALSE}
x <- c(2.23,3.45,1.87,2.11,7.33,18.34,19.23)
print(mean(x))
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Exercise #3

<font size="0.7">1 point possible (graded)</font>

Use a for loop to determine the value of $\sum_{i=1}^{25} i^2$

```{r 3, echo=FALSE}
for(i in 1:25) {
  x[i] <- i^2
}
print(sum(x))
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Exercise #4

<font size="0.7">1 point possible (graded)</font>

The `cars` dataset is available in base R. You can type `cars` to see it. Use the `class` function to determine what type of object is `cars`

```{r 4, , rows.print=50, cols.print=2, echo=FALSE}
cars
class(cars)
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Exercise #5

<font size="0.7">1 point possible (graded)</font>

How many rows does the `cars` object have?

```{r 5, echo=FALSE}
dim(cars)[1]
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Exercise #6

<font size="0.7">1 point possible (graded)</font>

What is the name of the second column of `cars`?

```{r 6, echo=FALSE}
names(cars)[2]
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Exercise #7

<font size="0.7">1 point possible (graded)</font>

The simplest way to extract the columns of a `matrix` or `data.frame` is using `[`. For example you can access the second column with `cars[,2]`. What is the average distance traveled in this dataset?

```{r 7, echo=FALSE}
mean(cars[,2])
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Exercise #8

<font size="0.7">1 point possible (graded)</font>

Familiarize yourself with the `which` function. What row of `cars` has a distance of 85?

```{r 8, echo=FALSE}
which(cars[,2]==85)
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
