---
title: "Introduction to dplyr"
author: "Maurizio La Rosa"
date: "15/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the data

```{r data}
mice_data <- read.delim("femaleMiceWeights.csv",header=TRUE,sep=",")
View(mice_data)
print(mice_data)
```

### Load the `dplyr` package

```{r dplyr}
library(dplyr)
```

### Command: `filter`

`filter` lets you filter certain rows, certain observations of the data set. Here i filter by the control group, i.e. the mice whose diet is not high fat. What I effectively get is a subset of the data set.

```{r filter, echo=FALSE}
controls <- filter(mice_data,Diet=="chow")
View(controls)
print(controls)
```


### Command: `select`

`select` lets you select certain columns, certain variables in the data set. I will use it to select the weights in the control group and then make a numerical vector of the selection, because many of the functions I'm going to use later make use of that.

```{r select, echo=FALSE}
controls_w1 <- select(controls,Bodyweight)
controls_w <- unlist(controls_w1)
print(controls_w)
```

### Using a `pipe` within `dplyr`

Making use of a `pipe` lets me perform the above steps in only one line of intuitive code. I perform the above steps for the high fat group.

```{r pipe, echo=FALSE}
treated_w <- filter(mice_data,Diet=="hf") %>% select(Bodyweight) %>% unlist
print(treated_w)
```

### Exploring the data

After having treated the data with filtering and selecting those of interest, I can perform data analysis and plotting

```{r analysis, echo=FALSE}
mean(controls_w)
mean(treated_w)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
