---
title: "edX - Statistics and R"
author: "Maurizio La Rosa"
date: "04/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
# Populations, parameters, and sample estimates Exercises
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

Load the mice_pheno data set.

```{r load the data, echo=FALSE}
mice <- read.delim("mice_pheno.csv",header=TRUE,sep=",")
mice <- na.omit(mice)
head(mice)
```

## Populations, parameters, and sample estimates Exercises #1

<font size="0.7">1 point possible (graded)</font>

Use `dplyr` to create a vector `x` with the body weight of all males on the control (`chow`) diet. What is this population's average?

```{r 1, echo=FALSE}
library(dplyr)
male_mice <- filter(mice,Sex=="M"&Diet=="chow") #filter male observations with chow diet
head(male_mice)
x <- select(male_mice,Bodyweight) %>% unlist() #select bodyweight column and unlist it
head(x)
mean(x)
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Populations, parameters, and sample estimates Exercises #2

<font size="0.7">1 point possible (graded)</font>

Now use the `rafalib` package and use the `popsd` function to compute the population standard deviation.

```{r 2, echo=FALSE}
#install.packages("rafalib")
library(rafalib)
popsd(x)
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Populations, parameters, and sample estimates Exercises #3

<font size="0.7">1 point possible (graded)</font>

Set the `seed` at 1. Take a random sample $X$ of size 25 from `x`. What is the sample average?

```{r 3, echo=FALSE}
set.seed(1)
X <- sample(x,25)
mean(X)
```

The result returned by sampling is different from the correct one, despite setting the seed at 1. This happens because the course was built when a previous version of R existed. Since R 3.6.0, the random number generator (RNG) used by the `sample` function has been changed. In order to use the old RNG one should first call:

`RNGkind(sample.kind = "Rounding")`

After that one can proceed normally. Hence:

```{r 3 bis, echo=FALSE}
RNGkind(sample.kind = "Rounding")
set.seed(1)
X <- sample(x,25)
mean(X)
```

Now this is the correct answer to Exercise 3.

<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Populations, parameters, and sample estimates Exercises #4

<font size="0.7">1 point possible (graded)</font>

Use `dplyr` to create a vector `y` with the body weight of all males on the high fat `hf`) diet. What is this population's average?

```{r 4, echo=FALSE}
library(dplyr)
y <- filter(mice,Sex=="M"&Diet=="hf") %>% select(Bodyweight) %>% unlist() #filter male observations with hf diet, select bodyweight column and unlist it
head(y)
mean(y)
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Populations, parameters, and sample estimates Exercises #5

<font size="0.7">1 point possible (graded)</font>

Now use the `rafalib` package and use the `popsd` function to compute the population standard deviation.

```{r 5, echo=FALSE}
library(rafalib)
popsd(y)
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Populations, parameters, and sample estimates Exercises #6

<font size="0.7">1 point possible (graded)</font>

Set the seed at 1. Take a random sample $Y$ of size 25 from `y.` What is the sample average?

```{r 6, echo=FALSE}
#Here I deal with the same problem as in Exercise 3.
RNGkind(sample.kind = "Rounding")
set.seed(1)
Y <- sample(y,25)
mean(Y)
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Populations, parameters, and sample estimates Exercises #7

<font size="0.7">1 point possible (graded)</font>

What is the difference in absolute value between $\bar{y}-\bar{x}$ and $\bar{Y}-\bar{X}$?

```{r 7, echo=FALSE}

abs((mean(y)-mean(x))-(mean(Y)-mean(X)))
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Populations, parameters, and sample estimates Exercises #8

<font size="0.7">1 point possible (graded)</font>

Repeat the above for females. Make sure to set the seed to 1 before each sample call. What is the difference in absolute value between $\bar{y}-\bar{x}$ and $\bar{Y}-\bar{X}$?

```{r 8, echo=FALSE}
x_ <- filter(mice,Sex=="F"&Diet=="chow") %>% select(Bodyweight) %>% unlist()
y_ <- filter(mice,Sex=="F"&Diet=="hf") %>% select(Bodyweight) %>% unlist()
RNGkind(sample.kind = "Rounding") #call old RNG
set.seed(1)
X_ <- sample(x_,25)
set.seed(1)
Y_ <- sample(y_,25)
abs((mean(y_)-mean(x_))-(mean(Y_)-mean(X_)))
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Populations, parameters, and sample estimates Exercises #9

<font size="0.7">1 point possible (graded)</font>

For the females, our sample estimates were closer to the population difference than with males. What is a possible explanation for this?

The population variance of the females is smaller than that of the males; thus, the sample variable has less variability.

<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
