---
title: "edX - Statistics and R"
author: "Maurizio La Rosa"
date: "26/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
# Probability Distributions
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

### Load the data

```{r data}
library(dplyr)
father_son <- read.delim("father-son.csv",header=TRUE,sep=",")
```

### Plot the distribution of the fathers' heights

Since it would be impossible or tedious or uncomfortable to have a look at 1078 observations at once, we rely on their representation by a histogram, which shows the occurrencies for an interval of realizations of heights.

```{r means, echo=FALSE}
hist(father_son$fheight)
```
From the histogram we can obtain useful information of the data. For example, we can compute the proportion of fathers whose heights fall below any given value. Given a specific value (say $a$), by using the histogram, we can compute pretty precisely the number of fathers who are shorter than height $a$ (and, by dividing them for the total number of fathers, we know their proportion).

$F(a)=Pr(Height≤a)$

In addition, given two values (say $a$ and $b$), using the histogram we can compute the proportion of fathers whose heights fall within that interval.
Given $b>a$ the values of the proportions of fathers whose heights are below $a$ and below $b$, it would suffice to subtract the second proportion from the first one to get the result:

$Pr(a≤Height≤b)=F(b)-F(a)$

### Plot the Cumulative Distribution Function

A function in the `stats` package allows us to compute the CDF of our values: `ecdf` (*empirical cumulative distribution function*), and we can consequently plot it. The CDF simplifies the computation of the above mentioned proportions because it adds up marginal proportions: this means that, if we pick any value $x$, the attached proportion ($f(x)$) returns the proportion of all values smaller than x $Pr(Height)≤x$.

```{r ecdf, echo=FALSE}
ecdf(father_son$fheight)
layout(matrix(c(1,1,1,1,1,1),nrow=3,byrow=TRUE))
plot(ecdf(father_son$fheight),main="",xlab="x=Height",ylab="F(x)")
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.