---
title: "edX - Statistics and R"
author: "Maurizio La Rosa"
date: "02/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
# Normal Distribution Exercises
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

Load the femaleControlsPopulation data set. Here x represents the weights for the entire population.

Using the same process as before (in Null Distribution Exercises), set the seed at 1, then using a for-loop take a random sample of 5 mice 1,000 times. Save these averages. After that, set the seed at 1, then using a for-loop take a random sample of 50 mice 1,000 times. Save these averages.

We recommend you review the video and read the section on normal distribution in the book before tackling these problems:

http://genomicsclass.github.io/book/pages/random_variables.html#normal_distribution

```{r load the data, echo=FALSE}
control_pop <- unlist(read.delim("femaleControlsPopulation.csv",header=TRUE,sep=","))
n <- 1000
set.seed(1)
samp_avg <- vector("numeric",n)
for(i in 1:n) {
  r_samp <- sample(control_pop,5)
  samp_avg[i] <- mean(r_samp)
}

set.seed(1)
samp_avg2 <- vector("numeric",n)
for(i in 1:n) {
  r_samp2 <- sample(control_pop,50)
  samp_avg2[i] <- mean(r_samp2)
}
```


## Normal Distribution Exercises #1

<font size="0.7">1 point possible (graded)</font>

Use a histogram to "look" at the distribution of averages we get with a sample size of 5 and a sample size of 50. How would you say they differ?

```{r 1, echo=FALSE}
par(mfrow=c(1,2))
hist(samp_avg,freq=FALSE)
hist(samp_avg2,freq=FALSE)
```

They both look roughly normal, but with a sample size of 50 the spread is smaller.
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Normal Distribution Exercises #2

<font size="0.7">1 point possible (graded)</font>

For the last set of averages, the ones obtained from a sample size of 50, what proportion are between 23 and 25?

```{r 2, echo=FALSE}
mean(samp_avg2<25)-mean(samp_avg2<23)
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Normal Distribution Exercises #3

<font size="0.7">1 point possible (graded)</font>

Now ask the same question of a normal distribution with average 23.9 and standard deviation 0.43.

```{r 3, echo=FALSE}
x <- rnorm(1000,23.9,0.43)
mean(x<25)-mean(x<23)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
