---
title: "edX - Statistics and R"
author: "Maurizio La Rosa"
date: "15/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
# dplyr Exercises
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

For these exercises, we will use a new dataset related to mammalian sleep. This link describes the data.
We are going to read in this data, then test your knowledge of they key dplyr functions select and filter. We are also going to review two different _classes_: data frames and vectors.

## dplyr Exercises #1

<font size="0.7">1 point possible (graded)</font>

Read in the `msleep_ggplot2.csv` file with the function `read.csv` and use the function `class` to determine what type of object is returned.

```{r 1, echo=FALSE}
data <- read.delim("msleep_ggplot2.csv",header=TRUE,sep=",")
View(data)
head(data)
class(data)
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## dplyr Exercises #2

<font size="0.7">1 point possible (graded)</font>

Now use the `filter` function to select only the primates. How many animals in the table are primates? Hint: the `nrow` function gives you the number of rows of a data frame or matrix.

```{r 2, echo=FALSE}
primates <- filter(data,order=="Primates")
nrow(primates)
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## dplyr Exercise #3

<font size="0.7">1 point possible (graded)</font>

What is the class of the object you obtain after subsetting the table to only include primates?

```{r 3, echo=FALSE}
class(primates)
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## dplyr Exercise #4

<font size="0.7">1 point possible (graded)</font>

Now use the `select` function to extract the sleep (total) for the primates. What class is this object? Hint: use `%>%` to pipe the results of the `filter` function to `select.`

```{r 4, echo=FALSE}
filter(data,order=="Primates") %>% select(sleep_total) %>% class
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## dplyr Exercise #5

<font size="0.7">1 point possible (graded)</font>

Now we want to calculate the average amount of sleep for primates (the average of the numbers computed above). One challenge is that the `mean` function requires a vector so, if we simply apply it to the output above, we get an error. Look at the help file for `unlist` and use it to compute the desired average.

```{r 5, echo=FALSE}
primates_sleep <- unlist(select(primates,sleep_total))
mean(primates_sleep)
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## dplyr Exercise #6

<font size="0.7">1 point possible (graded)</font>

For the last exercise, we could also use the dplyr `summarize` function. We have not introduced this function, but you can read the help file and repeat exercise 5, this time using just `filter` and `summarize` to get the answer.

```{r 6, echo=FALSE}
filter(data,order=="Primates") %>% summarise(mean=mean(sleep_total))
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.