---
title: "edX - Statistics and R"
author: "Maurizio La Rosa"
date: "23/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
# Random Variables Exercises
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

For these exercises, we will use the dataset of the entire mice control population. We load it and `unlist` it, in order to deal with a numerical vector.
`control_pop` represents the weights for the entire mice population.

```{r load the data, echo=FALSE}
control_pop <- unlist(read.delim("femaleControlsPopulation.csv",header=TRUE,sep=","))
head(control_pop)
```


## Random Variables Exercises #1

<font size="0.7">1 point possible (graded)</font>

What is the average of these weights?

```{r 1, echo=FALSE}
mean(control_pop)
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Random Variables Exercises #2

<font size="0.7">1 point possible (graded)</font>

After setting the seed at 1, `set.seed(1)` take a random sample of size 5. What is the absolute value (use `abs`) of the difference between the average of the sample and the average of all the values?

```{r 2, echo=FALSE}
set.seed(1)
sample_pop1 <- sample(control_pop,5)
abs(mean(sample_pop1)-mean(control_pop))
```
The result is wrong. \
However, setting the seed at 1 only returns this value (the accepted result is: 0.2706222, hence the sample mean should be around: 23.62276).
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Random Variables Exercise #3

<font size="0.7">1 point possible (graded)</font>

After setting the seed at 5, `set.seed(5)` take a random sample of size 5. What is the absolute value of the difference between the average of the sample and the average of all the values?

```{r 3, echo=FALSE}
set.seed(5)
sample_pop5 <- sample(control_pop,5)
abs(mean(sample_pop5)-mean(control_pop))
```

The result is wrong. \
However, setting the seed at 5 only returns this value (The accepted result is: 1.433378, hence the mean should be: 22.46)
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Random Variables Exercise #4

<font size="0.7">1 point possible (graded)</font>

Why are the answers from 2 and 3 different?`

Because we made a coding mistake. \
Because the average of the x is random. \
Because the average of the samples is a random variable. \
All of the above. \

```{r 4, echo=FALSE}
answ_opt <- c("Because we made a coding mistake","Because the average of the x is random","Because the average of the samples is a random variable","All of the above")
answ_opt[3] #my answer
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.