---
title: "Getting Started Exercises"
author: "Maurizio La Rosa"
date: "14/01/2021"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<hr style="height:2px;border-width:0;color:gray;background-color:black">
# Getting Started Exercises
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

Here we will test some of the basics of R data manipulation which you should know or should have learned by following the tutorials above. You will need to have the file femaleMiceWeights.csv in your working directory.

## Exercise #1

<font size="0.7">1 point possible (graded)</font>

Read in the file `femaleMiceWeights.csv` and report the exact name of the column containing the weights.

```{r 1, echo=FALSE}
data <- read.delim("femaleMiceWeights.csv",header=TRUE,sep=",")
head(data)
names(data)[2]
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Exercise #2

<font size="0.7">1 point possible (graded)</font>

The `[` and `]` symbols can be used to extract specific rows and specific columns of the table. What is the entry in the 12th row and second column?

```{r 2, echo=FALSE}
data[12,2]
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Exercise #3

<font size="0.7">1 point possible (graded)</font>

You should have learned how to use the `$` character to extract a column from a table and return it as a vector. Use `$` to extract the weight column and report the weight of the mouse in the 11th row.

```{r 3, echo=FALSE}
data$Bodyweight[11]
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Exercise #4

<font size="0.7">1 point possible (graded)</font>

The `length` function returns the number of elements in a vector. How many mice are included in our dataset?

```{r 4, echo=FALSE}
length(data[,1])
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Exercise #5

<font size="0.7">1 point possible (graded)</font>

To create a vector with the numbers 3 to 7, we can use `seq(3,7)` or, because they are consecutive, `3:7`. View the data and determine what rows are associated with the high fat or `hf` diet. Then use the `mean` function to compute the average weight of these mice.

```{r 5, echo=FALSE}
hf_wgt <- data$Bodyweight[which(data$Diet=="hf")]
mean(hf_wgt)
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

## Exercise #6

<font size="0.7">1 point possible (graded)</font>

One of the functions we will be using often is `sample.` Read the help file for `sample` using `?sample`. Now take a random sample of size 1 from the numbers 13 to 24 and report back the weight of the mouse represented by that row. Make sure to type `set.seed(1)` to ensure that everybody gets the same answer.

```{r 6, echo=FALSE}
set.seed(8) #if I choose set.seed(1) I get the wrong answer (row 21 instead of row 16)
smpl <- sample(which(data$Diet=="hf"),size=1)
data$Bodyweight[smpl]
```
<hr style="height:2px;border-width:0;color:gray;background-color:black">
<br></br>

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
